---
title: "basics"
author: "Juerg Schelldorfer"
date: "31 May 2019"
output:
  html_document: default
#   md_document:
#    variant: markdown_github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1 - What is R?

R is a computing environment that combines:

* a programming language called S, developed by John Chambers at Bell Labs, that implements the idea of programming with data (Chambers 1998),

* an extensive set of functions for classical and modern statistical data analysis and modeling,

* powerful numerical analysis tools for linear algebra, differential equations, and stochastics,

* graphics functions for visualizing data and model output,

* a modular and extensible structure that supports a vast array of optional **add-on packages**, and

* extensive help and documentation facilities.

* **free** and **open source**

* widely usedboth in academia and industry

* teaser:  http://shiny.rstudio.com/gallery


R is an open source software project, available for free download (R Core Team 2014a). Originally a research project in statistical computing (Ihaka and Gentleman 1996), it is now managed by a development team that includes a number of well-regarded statisticians, and is widely used by statistical researchers and working scientists as a platform for making new methods available to users.

R has been developed by statisticians and is hence very **convenient for actuaries**.

## 2 - What is RStudio?

Rtudio (https://www.rstudio.com/) is an integrated Development Environment (IDE) for R:

* like Microsoft Word, Excel, etc.
* built to help you write R code, run R code, and analyze data with R
* text editor, latex integration, debugging tool, version control
* Easy reporting via RShiny

To work with RStudio is one option to work with R. Other options are using Jupyter Notebooks (https://jupyter.org/).

RStudio consists of four different panes, each keeps track of separate information. 

* R Console
* R Scipt
* Plot
* Help files

See a short video on https://www.rstudio.com/products/RStudio/#Desktop

## 3 - Calculations

### R as a simple caclulator
```{r}
# Calculate 3 + 4
sqrt(2)
x <- 3
y <- x^2
x + y
sin(2*pi)
```

### Creating vectors
```{r}
c(1, 5, 80)
2:11
a <- c(1, 6, 10, 22, 7, 13)
mean(a)
sum(a)
```

### Creating matrices and data frames
```{r}
matrix()
m <- matrix(1:6, nrow=3, ncol=2, byrow = TRUE)
m
```

```{r}
data.frame()
df <- data.frame(Name = c("I","You","He"),Gender = as.factor(c(0,1,0)), Age = c(21,47,33))
df
```

### An **R statement** may consist of...

* an asignment stores the result of the calculation under temp_a and temp_b
```{r}
temp_a <- 3 * (4 + 2)
temp_b <- temp_a + 2.5
```
* a name of an object: display object
```{r}
temp_a
```

* a call to a function:  numerical or graphical result
```{r}
mean(c(temp_a,temp_b))
mn <- mean(c(temp_a,temp_b))
```

A function is called by its name followed by ().

### Assignment and Workspace

* Everything in R is an object and has a certain name like temp_a, mean, mn.
* R stores objects in your workspace
```{r}
temp_a <- 3 * (4 + 2)
```

* ATTENTION: Overwriting an object in R throughs no warning
```{r}
temp_a <- temp_b^2
temp_a
```

## 4 - Where to find help?

* Help about any function
```{r}
# ?foo
?lm
```
* If you have any question, google 'how do I...with R'.
+ huge community
+ already asked by somebody else
* Very useful and helpful Q&A website:  http://stackexchange.com/Cheat
* Sheet for Base R https://www.rstudio.com/resources/cheatsheets/
* R Reference Card https://cran.r-project.org/doc/contrib/Baggott-refcard-v2.pdf
* **Learning by doing** is particularly true for programming

## 5 - Data Import and Export

How do we get our data into R?

### Loading data from R (for training)
There are several packages containing claims data for Non-Life insurance.
```{r}
library(MASS)
#library(CASdatasets)
#library(insuranceData)
```

For example, laod the data "Insurance"" from the MASS package:
```{r}
data("Insurance")
?Insurance
head(Insurance)
```

### Loading data from file

Often, you have your data as a .csv file available. Chek *?read.table* for more information about this function.

```{r}
data_path <- "C:/Users/juerg/AktuarDataScience/Varia/2019_SSA/"
df <- read.table(paste(data_path,"dataCar.csv",sep=""), header = TRUE,sep=",")
# ?read.csv and many other specific loading packages available
```

Some useful functions to correctly load/save the data.

Let's set the working directory - the folder from which to operate (e.g.,save and load from).  Use:
```{r}
getwd()  ## print the current working directory
#setwd("yourpath")
```

Or alternatively in RStudio use 'Session -> Set Working Directory -> Choose Directory...'

Importing data in R is easy. Different ways depending on the format (csv, txt, xlsx, etc.).
Alternative:  use the 'Import Dataset' tool in RStudio (upper-rightpanel)

```{r}
data_path <- "C:/Users/juerg/AktuarDataScience/Varia/2019_SSA/"
df <- read.table(paste(data_path,"dataCar.csv",sep=""), header = TRUE,sep=",")
str(df)
```

To save or write data to a file:

```{r}
#write.table(df, file = "xy.txt", sep = " ")
```

where x is the data object to be stored an xy.txt.

Excel-files:  Use CSV
* > write.csv(...)
* > write.csv2(...)

## 6 - R Objects

Everything in R is an object, for example
* data frame:  most essential data structure in R
```{r}
str(Insurance)
```
* vector, e.g.  a column of the data set Insurance
```{r}
Age <- Insurance$Age
str(Age)
```

Goal is to look at or use a part of your object. To access only part of an object, use[ ]:
* for vectors: myvector[x]
* for two-dimensional objects, e.g.  data frames or matrices: mydata.frame[x, y]
Specify the indices by a vector (e.g.c(1, 2, 6)) and separate the indices of different dimensions by commas.

Let’s play around with the indexing of a data frame:  two-dimensionalobject!

```{r}
Insurance[ , ]
c(1, 3, 7)
1:10
Insurance[1:10 , ]
Insurance[-c( 1, 3, 7), ] # negative indices are excluded> d.sport[ , 2:3]
Insurance[c(1, 3, 6), 2:3]
```

## 7 - R Functions

Example function calls
```{r}
mean(Insurance$Claims)
quantile(Insurance$Claims)
quantile(Insurance$Claims, probs = c(0.75, 0.9))
```
Always check out the functions help function with ?mean and ?quantile.

Functions consist of mandatory and optional arguments:
mean(x, trim = 0, na.rm = FALSE, ...)
x:  mandatory argument
trim:  optional argument, default is 0
na.rm:  optional argument, default is FALSE

The arguments of a function have a defined order and each argument has its own unique name.
```{r}
mean(x = Insurance$Claims, na.rm = TRUE)
```
You can either use the names of the arguments, or place the values in the correct order (or a mix of both):
```{r}
mean(Insurance$Claims, ,TRUE)
```
Example functions with no mandatory arguments:  matrix(), vector(),array(), list()> ?matrix


## 8 - Useful functions

Useful functions (look for help by typing ?str):

* str()
* nrow() and ncol()
* dim()
* summary()
* apply()
* head() and tail()
see also R Reference card

## 9 - Packages

By default, R only provides a basic set of functions.  Additional functions(and datasets) are obtained by loading additional * add-on packages*

Install and load "MASS" package (https://cran.r-project.org/web/packages/MASS/MASS.pdf. There is always a pdf containing information about the package.
```{r}
#install.packages("MASS") # install onto computer once
require(MASS) # for every R session.
library(MASS)
```

Online resources:

* list of all packages:http://cran.r-project.org/web/packages/I

* by topic: http://cran.r-project.org/web/views/I

* ask Google

## 10 - Further reading

For continuation on that level, see
ftp://ess.r-project.org/users/sfs/RKurs/R.Intro/slides.pdf

Google's R Style Guide ("how to write good R code")
https://google.github.io/styleguide/Rguide.xml

RStudio Cheat Sheets (phantastic!)
https://www.rstudio.com/resources/cheatsheets/
